The impacts of park access on health outcomes: a spatial comprehensive approach

### 2018-12-11 manuscript.Rnw
Change to single-column for draft
Add spatial econometrics discussion

\documentclass[Afour,sageh,times]{sagej}
\def\journalname{Submitted to Environment and Planning B: Urban Analytics and City Science}

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

\begin{table*}[ht]
  \caption{Commonly cited benefits of parks and greenspaces}
  \label{tab:parkbenefits}
\begin{tabular}{lp{14cm}}
  \hline
Category & Summary of Benefits\\
\hline
Physical Health & Provide clean drinking water \citep{Benedict2006}\\
 & Foster increased physical activity\\
 & Promote faster healing in hospitals \citep{Akbari2001}\\
 & Reduce of heat-related mortality \\
 & Reduce incidence of cardiovascular-related mortality \\
 & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
 & Reduce incidence of childhood asthma \citep{Wolf2008}\\
 & Improve birth outcomes\\
\hline
Mental Health & Reduce stress and mental fatigue\\
  & Reduce aggression\\
  & Enhance emotional and cognitive development\\
  & Improve behavioral outcomes in youth \\
\hline
Social Health & Enhanced community aesthetics\\
 & Reduce crime\\
 & Increase social interaction \\
\hline
Economic Health & Provide of ecosystem services\\
 & Increase residential property values and municipal property tax revenues\\
 & Attract more shoppers and increase economic activity to commercial districts \\
\hline
\end{tabular}
\end{table*}

@@

$$ P(p | V_p) = \frac{\exp(V_p)}{\sum_{j \in J}\exp(V_j)}$$
where parks are differentiated from each other by their relative measurable
utilitie $V$. In principle, $V$ may include any measurable attributes of either
the choice maker or the park. In this study we use a linear formulation of

@@

In the absence of a park choice survey, we estimated likely values of the $\beta$
coefficients by iteratively searching for the values which produced the highest
model likelihood in *which model did we use to estimate likelihood?*. The search
was constrained by two assertions. First, we required that the coefficient on
size be positive and that on distance negative; all else equal, people will
prefer to use larger parks that are nearer to their residence. Second, we visually
inspected the resulting accessibility scores to ensure that the scores produced a
reasonably varied pattern of access throughout the metropolitan region.

@@

@@

\subsection{Spatial weight and model selection}

@@

>>>>

\documentclass[Crown,sageh,times]{sagej}
\def\journalname{Submitted to Environment and Planning B: Urban Analytics and City Science}

\usepackage{amsmath}
     \def\Var{{\rm Var}\,}
  	 \def\Cov{{\rm Cov}\,}
     \def\E{{\rm E}\,}
		 \def\Corr{{\rm Corr}\,}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\?}{\overset{?}{=}}

\begin{document}

@@

\begin{table*}[ht]
  \caption{Commonly cited benefits of parks and greenspaces}
  \label{tab:parkbenefits}
\begin{tabular}{lp{11cm}}
  \hline
Category & Summary of Benefits\\
\hline
Physical Health & Provide clean drinking water \citep{Benedict2006}\\
 & Foster increased physical activity\\
 & Promote faster healing in hospitals \citep{Akbari2001}\\
 & Reduce of heat-related mortality \\
 & Reduce incidence of cardiovascular-related mortality \\
 & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
 & Reduce incidence of childhood asthma \citep{Wolf2008}\\
 & Improve birth outcomes\\
\hline
Mental Health & Reduce stress and mental fatigue\\
  & Reduce aggression\\
  & Enhance emotional and cognitive development\\
  & Improve behavioral outcomes in youth \\
\hline
Social Health & Enhanced community aesthetics\\
 & Reduce crime\\
 & Increase social interaction \\
\hline
Economic Health & Provide of ecosystem services\\
 & Increase residential property values and municipal property tax revenues\\
 & Attract more shoppers and increase economic activity to commercial districts \\
\hline
\end{tabular}
\end{table*}

@@

$$ P(p | V_p) = \frac{\exp(V_p)}{\sum_{j \in J}\exp(V_j)}$$
where parks are differentiated from each other by their relative measurable
utilities $V$. In principle, $V$ may include any measurable attributes of either
the choice maker or the park. In this study we use a linear formulation of

@@

@@

We obtained data on aggregate health outcomes from the Centers for Disease
Control and Prevention (CDC) for census tracts in the New York City area. After
removing tracts with missing population information, we have Sexpr{nrow(tracts)}
complete cases.

@@

We also collected data on park geometry.

\subsection{Park Choice Logsum}
In the absence of a park choice survey, we estimated likely values of the $\beta$
coefficients by iteratively searching for the values which produced the highest
model likelihood in *which model did we use to estimate likelihood?*. The search
was constrained by two assertions. First, we required that the coefficient on
size be positive and that on distance negative; all else equal, people will
prefer to use larger parks that are nearer to their residence. Second, we visually
inspected the resulting accessibility scores to ensure that the scores produced a
reasonably varied pattern of access throughout the metropolitan region.

\subsection{Spatial dependence and correlation}

Using tracts as the unit of observation in the model allows us to consider the
availability of parks at a metropolitan level, but it also introduces the problem
spatial correlation of unobserved errors and spatial dependence in the data
generating process \citep{Cliff1970, Anselin1980}. Spatial correlation occurs
when spatially-distributed unobservable or missing variables (school quality,
neighborhood prestige, etc.) influence the modeled outcomes. Spatial dependence
occurs when the outcome {\em depends} on the outcome in neighboring areas; for
example, seeing neighbors exercising may encourage exercise. The relaxed spatial
Durbin model (SDM) proposed by \citet{Burridge1981} controls for both processes:
\begin{equation}\label{eq:sdm}
  \vect{y}= \rho W\vect{y} + X\vect{\beta} + WX\vect{\gamma} + \vect{\epsilon}
\end{equation}
where $\rho$ is the estimated spatial correlation between the outcome variables
$y_i, y_j, \ldots  \in \vect{y}$, $W$ is a matrix specifying the spatial
relationship between observations $i$ and $j$ in the dataset,
$\vect{\beta}$ is a vector of estimable coefficients relating the attributes $X$
of an observation to its outcome, and $\vect{\gamma}$ is a vector of
estimated coefficients relating the attributes of an observation's {\em neighbors}
to its outcome.

The researcher must assert $W$ {\em a priori}, though there are several
common specifications \citep{Dubin1998}. In this study we specify that tracts
with centroids located within 1.8 miles of each other are considered neighbors,
and the strength of the relationship is the inverse of the distance between
them:
\begin{equation}\label{eq:weightsrule}
W_{ij} = \begin{cases}
  {\frac{1/d_{ij}} {\sum_{k=1}^n( 1/d_{ik})}} & \text{for $d_{ij} \leq 1.8$ miles}\\
  0  & \text{for $d_{ij} > 1.8$ miles}
  \end{cases}
\end{equation}

@@

It may in some cases be more efficient to estimate a restricted spatial autoregressive
model, that controls for only correlation or dependence. \citet{tra_autoregression_paper}
illustrate a decision algorithm --- proposed by \citet{Florax2003} --- which uses
a likelihood ratio test of the SDM against a restricted spatial error model (SEM):
if the likelihood ratio test rejects that the models are equivalent, then the SDM
should be used.

A final point to note about spatial autoregressive models is that the partial
derivative of the model with respect to a single regressor $\vect{x_k} \in X$
does not equal the estimated coefficient $\beta_k$. Consequently, the estimated
coefficients do not represent the {\em effect} of a variable on the outcome. The
direct effects of an observation's attribute on its outcome, the indirect effects
of neighbors' attributes on an outcome, and the total effect of own and neighbors'
attributes can attained via Monte Carlo simulation. In this study we use the
\texttt{spdep} package for R \citep{Bivand2013} for model estimation and impact
simulation.
